---
title: "titanic_graphs"
author: "Lev Barbash"
date: "2023-08-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load}
library(tidyverse)
data <- read.csv("titanic.csv", sep=',')
data$Sex <- ifelse(data$Sex == 'male', 'Male', 'Female')
data$Pclass <- paste('Class', data$Pclass)
data$Survived <- ifelse(data$Survived == 1, 'Yes', 'No')
```

## 1

Generate a series of bar charts to describe the gender, ticket class and survival of the passengers onboard.

```{r bar_charts_total}
library(ggplot2)
library(gridExtra)
columns <- c('Sex', 'Pclass', 'Survived')
plots <- list()
for (col in columns) {
  temp_sers <- table(data[[col]])
  
  p <- ggplot(data.frame(x = names(temp_sers), y = as.numeric(temp_sers)),
              aes(x = x, y = y)) +
    geom_bar(stat = 'identity', fill = 'steelblue') +
    labs(title = col) +
    ylim(0, 600) +
    theme_minimal()
  
  plots[[col]] <- p
}
grid.arrange(plots[[1]], plots[[2]], plots[[3]], ncol = 3)
```

## 2

Generate a histogram for the passengers’ age. Furthermore, describe the passengers’ age using the following two boxplots: age per ticket class and age based on survival.

```{r hist_age}
data <- na.omit(data)
p <- ggplot(data, aes(x = Age)) +
  geom_histogram(binwidth = 5, fill = 'steelblue', color = 'black') +
  labs(title = "Passengers' age distribution") +
  theme_minimal()

print(p)
```

```{r box_1}
p <- ggplot(data, aes(x = Pclass, y = Age)) +
  geom_boxplot(fill = 'steelblue', color = 'black') +
  scale_x_discrete(labels = c('Class 1', 'Class 2', 'Class 3')) +
  labs(title = 'Age per ticket class') +
  theme_minimal()

print(p)
```

```{r box_2}
p <- ggplot(data, aes(x = Survived, y = Age)) +
  geom_boxplot(fill = 'steelblue', color = 'black') +
  scale_x_discrete(labels = c('No', 'Yes')) +
  labs(title = 'Age based on survival') +
  theme_minimal()

print(p)
```

## 3

Generate a histogram for the travel fare and a table showing the number of people who did not pay – you may want to check on Google why a handful of people was on board for free!

```{r hist_fare}
p1 <- ggplot(data, aes(x = Fare)) +
  geom_histogram(binwidth = 10, fill = 'steelblue', color = 'black') +
  labs(title = "Passengers' fare distribution") +
  theme_minimal()

p2 <- ggplot(data[data$Fare <= 150, ], aes(x = Fare)) +
  geom_histogram(binwidth = 10, fill = 'steelblue', color = 'black') +
  labs(title = "(Fare <= 150)") +
  theme_minimal()

p3 <- ggplot(data[data$Fare <= 50, ], aes(x = Fare)) +
  geom_histogram(binwidth = 10, fill = 'steelblue', color = 'black') +
  labs(title = "(Fare <= 50)") +
  theme_minimal()

grid.arrange(p1, p2, p3, ncol = 3)
```

```{r free_fare_table}
print(data[data$Fare == 0, ])
```

## 4

A chart of your choice to describe the family size per ticket class

```{r pie_chart}
passenger_classes <- c('Class 1', 'Class 2', 'Class 3')

plots <- list()

for (cl in passenger_classes) {
  temp_sizes <- table(data[data$Pclass == cl, 'SibSp'])
  
  p <- ggplot(data.frame(SibSp = names(temp_sizes), Count = as.numeric(temp_sizes)),
              aes(x = "", y = Count, fill = SibSp)) +
    geom_bar(stat = "identity", width = 1) +
    geom_text(aes(label = paste0(round(Count/sum(Count) * 100, 1), "%")), position = position_stack(vjust = 0.5)) +
    coord_polar(theta = "y") +
    labs(title = cl) +
    theme_void() +
    theme(legend.position = "bottom")
  
  plots[[cl]] <- p
}

grid.arrange(plots[[1]], plots[[2]], plots[[3]], ncol = 3)
```

## 5

A series of stacked bar charts to show the how survival differs for different gender and ticket class

```{r surv_gend}
survival_counts <- data %>%
  filter(Survived == 'No' | Survived == 'Yes') %>%
  group_by(Sex, Survived) %>%
  tally()

p <- ggplot(survival_counts, aes(x = Sex, y = n, fill = Survived)) +
  geom_bar(stat = 'identity', position = 'stack') +
  labs(title = 'Survival Rate for Different Genders') +
  scale_fill_manual(values = c('No' = 'red', 'Yes' = 'green')) +
  theme_minimal()

print(p)
```

```{r surv_class}
survival_counts <- data %>%
  filter(Survived == 'No' | Survived == 'Yes') %>%
  group_by(Pclass, Survived) %>%
  tally()

# Create a stacked bar plot
p <- ggplot(survival_counts, aes(x = Pclass, y = n, fill = Survived)) +
  geom_bar(stat = 'identity', position = 'stack') +
  labs(title = 'Survival Rate for Different Ticket Classes') +
  scale_fill_manual(values = c('No' = 'red', 'Yes' = 'green')) +
  theme_minimal()

print(p)
```

## 6

A violin chart describing how survival related to age and gender

```{r violin_gend}
p <- ggplot(data, aes(x = Sex, y = Age, fill = Survived)) +
  geom_violin() +
  labs(title = 'Survival related to age and gender') +
  scale_fill_manual(values = c('No' = 'red', 'Yes' = 'green')) +
  theme_minimal()

print(p)
```

```{r violin_class}
p <- ggplot(data, aes(x = Pclass, y = Age, fill = Survived)) +
  geom_violin() +
  labs(title = 'Survival related to age and ticket class') +
  scale_fill_manual(values = c('No' = 'red', 'Yes' = 'green')) +
  theme_minimal()

print(p)
```